---
title: "Prediction"
author: "Alex Baur"
date: "March 22, 2016"
output: html_document
---

Will work off the Milestone1.RMD file.

Convert Vcorpus to Pcorpus for easy use and to keep out of RAM?
Need 4-gram for the final prediction in the case of a 3-input?

```{r}
setwd("C:/Users/alexb/Desktop/Coursera/Capstone/final/en_US")
load("C:/Users/alexb/Desktop/Coursera/Capstone/final/en_US/.RData")
```


## Four-gram for completion

```{r}
corpgram4 <- function(x) NGramTokenizer(x, Weka_control(min = 4, max = 4))

corpout4 <- TermDocumentMatrix(corpsub, control=list(tokenize = corpgram4))

fourgram <- rollup(corpout4, 2, na.rm=TRUE, FUN = sum)

fourgram <- data.frame(as.matrix(fourgram))

foursort <- fourgram[order(-fourgram[,"X1"]), , drop=FALSE]

foursub <- foursort[1:20, , drop=FALSE]

foursub$X2 <- rownames(foursub)

colnames(foursub) <- c("Frequency", "Phrase")
```

## Four-gram plot
```{r}
fourplot <- ggplot(foursub, aes(x=Phrase, y=Frequency))+geom_bar(stat="Identity", aes(reorder(Phrase,-Frequency),Frequency), fill="mediumorchid3", colour = "black")+geom_text(aes(label=Frequency), vjust = -0.25) + theme(axis.text.x = element_text(angle = 90, size=12, vjust = 0.35))

fourplot
```


## Assign column with row names for the grams, allows sorting
```{r}
onegram$X2 <- rownames(onegram)
twogram$X2 <- rownames(twogram)
threegram$X2 <- rownames(threegram)
fourgram$X2 <- rownames(fourgram)
```

## Then sort the grams (should have been done in these steps in the milestone)
```{r}
onesort <- onegram[order(-onegram[,"X1"]), , drop=FALSE]

twosort <- twogram[order(-twogram[,"X1"]), , drop=FALSE]

threesort <- threegram[order(-threegram[,"X1"]), , drop=FALSE]

foursort <- fourgram[order(-fourgram[,"X1"]), , drop=FALSE]

## Column Naming
colnames(onesort) <- c("Frequency", "Phrase")
colnames(twosort) <- c("Frequency", "Phrase")
colnames(threesort) <- c("Frequency", "Phrase")
colnames(foursort) <- c("Frequency", "Phrase")
```

## Remove sparse entries (freq=1)
```{r}
onesort <- onesort[!rowSums(onesort==1),]
twosort <- twosort[!rowSums(twosort==1),]
threesort <- threesort[!rowSums(threesort==1),]
foursort <- foursort[!rowSums(foursort==1),]
```

## Set Probabilities for most common words in case input does not match ngrams
```{r}
    onesub <-  onesort[1:25, , drop=FALSE]
    freqtot <- sum(onesub$Frequency)
    onesub$Probability <- (onesub$Frequency)/freqtot
```

## Word Prediction Function
```{r}
wordpred <- function(input =" "){
    suppressWarnings(require(tm))
    suppressWarnings(require(stringr))
    input <- PlainTextDocument(input)
    input <- tolower(input)
    input <- removeNumbers(input)
    input <- removePunctuation(input)
    input <- stripWhitespace(input)

    if (input == " "){stop("Invalid Input: Please enter a word or phrase")}
    
    
    alphacheck <- grepl("[a-zA-Z]", input)
    if (alphacheck == FALSE){stop("Invalid Input: Please enter a word or phrase")}
    
    f <- function(input){
        inputsep <- unlist(strsplit(input, " "))
        inputsep <- rev(inputsep)
        lengthsep <- length(inputsep)
        
        if (lengthsep>=3){
            inputsub <- inputsep[1:3]
            input2 <- rev(inputsub)
            input2 <- paste(input2[1], input2[2], input2[3], sep =" ")
            input2
        }
        
        if (lengthsep == 2){
            inputsub <- inputsep[1:2]
            input2 <- rev(inputsub)
            input2 <- paste(input2[1], input2[2], sep =" ")
            input2
        }
                
        if (lengthsep == 1){
            input2 <- inputsep
            input2
        }
    input2<-input2    
    }
    
    g <- function(input2){
        input2sep <- as.character(unlist(strsplit(input2, " ")))
        length2sep <- length(input2sep)
        
        if(length2sep == 3){
            threesearch <- grepl(paste0("^",input2,"\\s"), foursort$Phrase)
            threefind <- foursort[threesearch, ]
            threefind <- threefind[1, 2]
            threeout <- word(threefind, -1)
            out <- threeout
            out
            
            threecheck <- all(threesearch==FALSE)
            
            if (threecheck==TRUE){
                input3 <- unlist(strsplit(input2, " "))
                input3 <- input3[2:3]
                input3 <- paste(input3[1], input3[2], sep = " ")
                twobumpsearch <- grepl(paste0("^",input3,"\\s"), threesort$Phrase)
                twobumpfind <- threesort[twobumpsearch, ]
                twobumpfind <- twobumpfind[1, 2]
                twobumpout <- word(twobumpfind, -1)
                outbump2 <- twobumpout
                outbump2
                
                twobumpcheck <- all(twobumpsearch==FALSE)
                
                if (twobumpcheck==TRUE){
                    input4 <- unlist(strsplit(input2, " "))
                    input4 <- input4[3]
                    onebumpsearch <- grepl(paste0("^",input4,"\\s"), twosort$Phrase)
                    onebumpfind <- twosort[onebumpsearch, ]
                    onebumpfind <- onebumpfind[1, 2]
                    onebumpout <- word(onebumpfind, -1)
                    outbump1 <- onebumpout
                    outbump1
                } else {return(outbump2)}
            } else {return(out)}
        }
            
        if (length2sep == 2){
            twosearch <- grepl(paste0("^",input2,"\\s"), threesort$Phrase)
            twofind <- threesort[twosearch, ]
            twofind <- twofind[1, 2]
            twoout <- word(twofind, -1)
            out <- twoout
            out
                
                twocheck <- all(twosearch==FALSE)
                
                if (twocheck==TRUE){
                    input4 <- unlist(strsplit(input2, " "))
                    input4 <- input4[2]
                    onebumpsearch <- grepl(paste0("^",input4,"\\s"), twosort$Phrase)
                    onebumpfind <- twosort[onebumpsearch, ]
                    onebumpfind <- onebumpfind[1, 2]
                    onebumpout <- word(onebumpfind, -1)
                    outbump1 <- onebumpout
                    outbump1
                } else {return(out)}
        }
        if (length2sep == 1){
                onesearch <- grepl(paste0("^",input2,"\\s"), twosort$Phrase)
                onefind <- twosort[onesearch, ]
                onefind <- onefind[1, 2]
                oneout <- word(onefind, -1)
                out <- oneout
                out
        }
    out    
    }
    
    input2 <- f(input)
    input2 <- as.character(input2)
    out <- g(input2)
    
    testna <- is.na(out)
    
    if (testna==TRUE){
        mostfreqout <- sample(onesub$Phrase, prob=onesub$Probability)
        
        repcheck <- word(input2, -1)==mostfreqout[1]
        
        if (repcheck==TRUE){
            mostfreqout[2]
        } else {return(mostfreqout[1])}
    } else {return(out)}
}
```

Implemented using:
-Katz-Backoff (going to lower order ngrams if higher order not found)
-Markov chains (word prediction dependent on only the last n words in a sentence)

## In progress
```{r}
        if (testna==TRUE){
        mostfreqout <- sample(onesub$Phrase, prob=onesub$Probability)
        mostfreqout[1]
    } else {out}
        
    if (testnull==TRUE){
        mostfreqout <- sample(onesub$Phrase, prob=onesub$Probability)
        mostfreqout[1]
    } else {out}
    
    
        out <- na.omit(out)
    out <- lapply(out, as.character)
    
        numcheck <- all(is.numeric(input))
    if (numcheck == TRUE){stop("Invalid Input: Please enter a word or phrase")}
    
            testna <- suppressWarnings(is.na(out))
    testnull <- suppressWarnings(is.null(out))
        
    if ((testna = TRUE)){
        mostfreqout <- sample(onesub$Phrase, prob=onesub$Probability)
        mostfreqout[1]
        } 
    if (((testnull=TRUE))){
        mostfreqout <- sample(onesub$Phrase, prob=onesub$Probability)
        mostfreqout[1]
    } else {out}
    
    
    
        testna <- is.na(out)
        testnull <- is.null(out)
        if ((testna = FALSE)|(testnull=FALSE)){
            mostfreqout <- sample(onesub$Phrase, prob=onesub$Probability)
            mostfreqout[1]
        }
        else {out}
    }
    h(out)
    
    outna <- na.omit(out)
    out <- lapply(outna, as.character)
    out
```


## Graveyard
```{r}
    # out <- data.frame(out)
    # out <- na.omit(out)
    # out <- lapply(out, as.character)
    # out
    # out
    
    if (input != [:alnum:]){return("Invalid Input: Please enter a word or phrase")}
    if (input != [:lower:]){return("Invalid Input")}

    library(stringr)
    ysearch <- grep(paste0("^",y,"\\s"), foursort$Phrase)
    yfind <- threesort[ysearch, ]
    yfind <- yfind[1:3, 2]
    yout <- word(yfind, -1)
    yout

    else
        if (lengthsep == 0 | lengthsep == " "){
            return("Invalid input")
        }

## Ex
Corpus[[1]]$content[1]
"I like dogs"

a<-unlist(strsplit(Corpus[[1]]$content[1]))
a
"I" "like" "dogs"

a[1]<-"<s>"
a
"<s>" "I" "like" "dogs"

Corpus[[1]]$content[1]<-paste(a)
Corpus[[1]]$content[1]
"<s> I like dogs"

## Ex
f <- function(query) {
require(tau)
trigrams <- sort(textcnt(rep(tolower(names(queryHistoryTab)), queryHistoryTab), method = "string", n = length(scan(text = query, what = "character", quiet = TRUE)) + 1))
query <- tolower(query)
idx <- which(substr(names(trigrams), 0, nchar(query)) == query)
res <- head(names(sort(trigrams[idx], decreasing = TRUE)), n)
res <- substr(res, nchar(query) + 2, nchar(res))
return(res)
}

f(c("Can of beer" = 3, "can of Soda" = 2, "A can of water" = 1, "Buy me a can of soda, please" = 2), "Can of")

# [1] "soda" "beer"
```

## Quiz 1
- The guy in front of me just bought a pound of bacon, a bouquet, and a case of         beer
- You're the reason why I smile everyday. Can you follow me please? It would mean the       world
- Hey sunshine, can you follow me and make me the       happiest
- Very early observations on the Bills game: Offense still struggling but the       defense
- Go on a romantic date at the      beach
- Well I'm pretty sure my granny has some old bagpipes in her garage I'll dust them off and be on my    way
- Ohhhhh #PointBreak is on tomorrow. Love that film and haven't seen it in quite some   time
- After the ice bucket challenge Louis will push his long wet hair out of his eyes with his little      fingers
- Be grateful for the good times and keep the faith during the      bad
- If this isn't the cutest thing you've ever seen, then you must be         insane

## Quiz 2
-When you breathe, I want to be the air for you. I'll be there for you, I'd live and I'd    die
-Guy at my table's wife got up to go to the bathroom and I asked about dessert and he started telling me about his     marital
-I'd give anything to see arctic monkeys this   weekend
-Talking to your mom has the same effect as a hug and helps reduce your     stress
-When you were in Holland you were like 1 inch away from me but you hadn't time to take a   picture
-I'd just like all of these questions answered, a presentation of evidence, and a jury to settle the    matter
-I can't deal with unsymetrical things. I can't even hold an uneven number of bags of groceries in each     hand
-Every inch of you is perfect from the bottom to the    top
-I am thankful my childhood was filled with imagination and bruises from playing     outside
-I like how the same people are in almost all of Adam Sandler's     movies

For application:
```{r}
All I do is put my data file in the same directory as ui.R and server.R, put the following in server.R:

load(file="filename.rdata")
```

